### 2020-04-06 paper2_choice of fig.Rmd
prep for final figures

---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

# Figures specification
* titled
* cited in numerical order in the text using Arabic numbers. 
* no additional charges for color. 
* submitted as individual files
* labeled with the Corresponding Author name, the appropriate figure number, and  orientation (e.g., “top”)
* Figures may be submitted as multipart panels.
*  *.tif* or .eps 
* final publication size; 1-column width or, if necessary, 1½ column width. The * 2-column width should not be used unless necessary.
* 1.0 column = 3.2 Inches, 8.2 Centimeters, 19.25 Picas
* 1.5 column = 4.5 Inches, 11.4 Centimeters, 27.00 Picas
* 2.0 column = 5.8 Inches, 14.8 Centimeters, 35.00 Picas
* height of all figures <= to 9.6inches / 24.5 cm / 58 picas.
* Calibri font sub sans serif Arial
* Multipanel labeled using uppercase 12-point Calibri Bold. 
* minimum text size of 6 points
* bold enough to be easily read after reduction, as should all symbols used in the * figure. Line or bar graphs or flow charts with text should be created in black and * white, not shades of gray, which are difficult to reproduce in even tones. If more * than two sets of data are represented, use of fill patterns or colors (not gray) is * suggested to present the data clearly.
* Minimum resolution is 300 dpi for color and grayscale figures, 600 dpi for * combination halftones, and 1000 dpi for line art.

```{r pkg}
library(tidyverse)
library(RColorBrewer)
library(readxl)
library(kableExtra)
library(tiff)
library(grid)
```

```{r}

# palette
blues <- brewer.pal(9,"Blues")
greens <- brewer.pal(9,"Greens")
excel_pal <- c("#CEDBE6", "#3494BA", "#58B6C0", "#75BDA7", "#84ACB6", "#2683C6", "#373545")
greys <- gray.colors(6, start = 0, end = 1)

# figure saving settings
units = "in"
fig_w <- 3.2
fig_w_1.5 <- 4.5
fig_h <- 4.5
fig_h_1.5 <- 4.5 * 1.5
dpi <- 300
device <- "tiff"

```

```{r import}
file <- here::here("data", "Meta-Analysis Data Collection.xlsx")

# added a row for column names more useful
# # skipped first two rows of of old col names

dat <- read_excel(file, sheet = "Data")

```

## Other palettes

All colour blind-friendly.

```{r}
display.brewer.all(colorblindFriendly = TRUE)
```
```{r fig.height=.2, fig.width=6}
df <- data.frame(x = 1:6, y = 1)
pal <- viridisLite::viridis(6)
df %>% ggplot(aes(x = factor(x), y = y, fill = factor(x))) +
  geom_col(fill = pal) +
  theme_void()
pal <- viridisLite::viridis(6, option = "A")
df %>% ggplot(aes(x = factor(x), y = y, fill = factor(x))) +
  geom_col(fill = pal) +
  theme_void()
pal <- viridisLite::viridis(6, option = "B")
df %>% ggplot(aes(x = factor(x), y = y, fill = factor(x))) +
  geom_col(fill = pal) +
  theme_void()
pal <- viridisLite::viridis(6, option = "C")
df %>% ggplot(aes(x = factor(x), y = y, fill = factor(x))) +
  geom_col(fill = pal) +
  theme_void()
```

Note: there is a 85th line containing only a note in the `NOTES` variable - presumably it belongs elsewhere and was excluded. The note was "compares different papers on bone marrow MSCs, doesn't say anything about prep of cells"

```{r tidy-data}
# recode country so those with one paper are
# "Other"
dat <- dat %>%
  mutate(Country2 = recode_factor(Country,
                                Brazil = "Brazil",
                                China = "China",
                                Denmark = "Denmark",
                                Egypt = "Egypt",
                                India = "India",
                                Japan = "Japan",
                                Korea = "Korea",
                                Netherlands = "Netherlands",
                                Poland = "Poland",
                                `Rep Korea` = "Rep Korea",
                                Russia = "Russia",
                                Spain = "Spain",
                                Taiwan = "Taiwan",
                                USA = "USA",
                                .default = "Other"))

dat <- dat %>%
  mutate(Phase_condensed = recode_factor(Phase,
                               I = "I",
                               `I/IIa` = "I",
                                II = "II",
                                IIa = "II",
                               IIb = "II",
                               III = "III"))

# test results - there is no "t" in the data file and there are more
# blanks values than t's in draft paper fig 6
dat <- dat %>%
  mutate(test_results = recode_factor(`Test Results?`,
                                       n = "Average value",
                                       s = "Stated as 'standard phenotype'",
                                       o = "Not discussed",
                                       y = "Phenotypic markers reported",
                                       t = "Tests done but not reported"))
dat$test_results <- fct_relevel(dat$test_results,
                              "Phenotypic markers reported", 
                              "Average value",
                              "Tests done but not reported",
                              "Stated as 'standard phenotype'",
                              "Not discussed")

# MOA - recode NS to Not stated
dat <- dat %>%
  mutate(MOA = recode_factor(MOA,
                             NS = "Not stated",
                             paracrine = "Paracrine",
                             immune = "Immune",
                             differentiation = "Differentiation",
                             multi = "Muliple"))

dat$MOA <- fct_relevel(dat$MOA,
                       "Paracrine",
                       "Immune",
                       "Differentiation",
                       "Muliple",
                       "Not stated")

# recode stem/stromal
dat <- dat %>%
  mutate(stem_stromal = recode_factor(`Stem/Stromal`,
                             `multipotent stromal` = "Multipotent Stromal",
                             other = "Other",
                             stem = "Stem",
                             stromal = "Stromal"))

dat$stem_stromal <- fct_relevel(dat$stem_stromal,
                                "Stromal",
                                "Stem",
                                "Multipotent Stromal",
                                "Other")

# Recode O, A and C into the 5 categories that appar 
dat <- dat %>%
  mutate(potency = ifelse(O == "y" &
                            A == "y" &
                            C == "y",
                          "Demonstrated tripotency",
                          ifelse(O == "n" &
                                   A == "n" &
                                   C == "n",
                                 "Not Stated",
                                 ifelse(O == "y" &
                                          A == "y" &
                                          C == "n",
                                        "Demonstrated bipotency (OA)",
                                        ifelse(O == "m"
                                               & A == "m" &
                                                 C == "n",
                                               "Claimed bipotency (OA)",
                                               ifelse(O == "m" &
                                                  A == "m" &
                                                  C == "m",
                                                  "Claimed tripotency", NA))))))

dat$potency <- fct_relevel(dat$potency,
                                "Demonstrated tripotency",
                                "Demonstrated bipotency (OA)",
                                "Claimed tripotency",
                                "Claimed bipotency (OA)",
                                "Not Stated")

######### code for "Data Tab copy v1.csv"
# # column indices for markers
# firstmarker <- which(colnames(dat) == "CD73")
# lastmarker <- which(colnames(dat) == "SSEA.4")
# # vector of marker names
# markers <- names(dat[firstmarker:lastmarker])
# 
# datnum <- dat %>% 
#   select(markers) %>% 
#   mutate_all(as.numeric) 
# 
# # add the non marker columns
# notmarkers <- which(!(colnames(dat) %in% markers))
# datnum <- cbind(dat[notmarkers], datnum)
# 
# 
# dat$Stem.Stromal <- factor(dat$Stem.Stromal)
# dat$Year <- factor(dat$Year)
# datnum$Stem.Stromal <- factor(datnum$Stem.Stromal)
# datnum$Year <- factor(datnum$Year)

```
# Figures
Note: tifs are of high quality than appear in this doc

## Fig 1 - Article search funnel

Funnel plot - could be adapted to include the reviews adjacent and aligned with the articles plot.

> to do: Add the Reviews diagram

```{r data-prep-fig1}
# create data
# short names for proof of method
orig <- data.frame(step_names = c("1.mesenchymal",
                                  "2.article",
                                  "3.cell tissue etc",
                                  "4.clinical etc",
                                  "5.not biotech etc",
                                  "6.exc conf etc"),
                   n = c(65706,
                         47031,
                         7814,
                         5502,
                         2055,
                         1986))

orig$step_names <- fct_relevel(orig$step_names,
                               "mesenchymal",
                               "article",
                               "cell tissue etc",
                               "clinical etc",
                               "not biotech etc",
                               "exc conf etc")

# prep polygons that connect
to_poly_orig <- orig %>%
  arrange(n) %>%
  mutate(
    group = row_number(),
    x1 = 0 + ((1 - n)/2),
    x4 = n + ((1 - n)/2),
    x2 = lag(x1), 
    x3 = lag(x4)) %>%
  mutate(
    x2 = ifelse(is.na(x2), x1, x2),
    x3 = ifelse(is.na(x3), x4, x3)
  ) %>%
  gather(xpoint, x, -step_names,-group, -n) %>%
  arrange(group, xpoint) %>%
  mutate(y = ifelse(xpoint %in% c("x2", "x3"), group, group + 1))

# EXAMPLE
# Create the placement of the stages in the plot,
#  by just adding a 0.5 to each row number

labels <- orig %>%
  arrange(n) %>%
  mutate(y = row_number() + 0.5)

# modify the labels
labels$step_names2 <- c( "EXCLUDE conference proceedings,\nretracted papers",
                         "EXCLUDE CATEGORIES (NOT\nbiotechnology applied microbiology oncology\nhaematology, engineering biomedical)",
                         "REFINE BY SOURCE TITLES\n(EXCLUDE NOT clinical/\ntranslation/regenerative medicine)",
                         "REFINE BY CATEGORY\ncell tissue engineering\nand transplantation",
                         "REFINE BY DOCUMENT TYPE\narticle",
                         "mesenchymal stem cells OR\nmesenchymal stromal cells OR\nMSC AND characterisation AND clinical")
```

```{r fig1-blues}
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               color = "black") +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 2,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  labs(title = "Wilson, Alison, Figure 1. TOP" ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
    scale_fill_manual(values = blues)

ggsave("figs/fig_1_Article_search_funnel_blues.tif", 
       plot = article_funnel, 
       device = device,
       width = fig_w, 
       height = 3.0,
       units = units,
       dpi = dpi)
```

```{r}
grid.raster(readTIFF("figs/fig_1_Article_search_funnel_blues.tif"))
```

```{r fig1-excel}
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               color = "black") +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 2,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  labs(title = "Wilson, Alison, Figure 1. TOP") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
    scale_fill_manual(values = excel_pal)

ggsave("figs/fig_1_Article_search_funnel_excel.tif", 
       plot = article_funnel, 
       device = device,
       width = fig_w, 
       height = 3.0,
       units = units,
       dpi = dpi)
```

```{r}
grid.raster(readTIFF("figs/fig_1_Article_search_funnel_excel.tif"))
```

```{r fig1-greys}
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               color = "black") +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 2,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  labs(title = "Wilson, Alison, Figure 1. TOP" ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
    scale_fill_manual(values = rev(greys))

ggsave("figs/fig_1_Article_search_funnel_greys.tif", 
       plot = article_funnel, 
       device = device,
       width = fig_w, 
       height = 3.0,
       units = units,
       dpi = dpi)
```

```{r}
grid.raster(readTIFF("figs/fig_1_Article_search_funnel_greys.tif"))
```

## Figure 3 Articles by country

```{r country-prep}
summary_country <- dat %>%
  group_by(Country) %>% 
  summarise(n = length(Country))

```

```{r fig3-country-bar-blues}
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(Country, n), y = n)) +
  geom_bar(stat = "identity", fill = blues[4],
           colour = "black") +
  xlab("") +
  scale_y_continuous(name = "Number of articles", 
                     breaks = seq(0, 16, 1),
                     minor_breaks = seq(0, 16, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 3. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 6, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()
ggsave("figs/fig_3_Article_country_blues.tif", 
       plot = country_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_3_Article_country_blues.tif"))
```

```{r fig3-country-bar-excel}
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(Country, n), y = n)) +
  geom_bar(stat = "identity", fill = excel_pal[3],
           colour = "black") +
  xlab("") +
  scale_y_continuous(name = "Number of articles", 
                     breaks = seq(0, 16, 1),
                     minor_breaks = seq(0, 16, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 3. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 6, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()
ggsave("figs/fig_3_Article_country_excel.tif", 
       plot = country_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_3_Article_country_excel.tif"))
```

```{r fig3-country-bar-greys}
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(Country, n), y = n)) +
  geom_bar(stat = "identity", fill = "white",
           colour = "black") +
  xlab("") +
  scale_y_continuous(name = "Number of articles", 
                     breaks = seq(0, 16, 1),
                     minor_breaks = seq(0, 16, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 3. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 6, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 6, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()
ggsave("figs/fig_3_Article_country_greys.tif", 
       plot = country_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_3_Article_country_greys.tif"))
```

## Figure 4 Articles by Phase

### Six phase version

```{r phase-prep}
summary_phase <- dat %>%
  group_by(Phase) %>% 
  summarise(n = length(Phase),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
summary_phase <- summary_phase %>%
  mutate(Phase_condensed = recode_factor(Phase,
                               I = "I",
                               `I/IIa` = "I",
                                II = "II",
                                IIa = "II",
                               IIb = "II",
                               III = "III"))

summary_phase_condensed <- dat %>%
  group_by(Phase_condensed) %>% 
  summarise(n = length(Phase_condensed),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

```

```{r fig4-phase-blues}
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(Phase,n), y = n)) +
  geom_bar(stat = "identity", fill = blues[4], colour = "black") +
  xlab("") +
  geom_text(aes(y = 10, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 30, 5),
                     minor_breaks = seq(0, 30, 1),
                     limits = c(0, 30),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase_blues.tif", 
       plot = phase_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_4_Article_phase_blues.tif"))
```

```{r fig4-phase-excel}
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(Phase,n), y = n)) +
  geom_bar(stat = "identity", fill = excel_pal[3], colour = "black") +
  xlab("") +
  geom_text(aes(y = 10, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 30, 5),
                     minor_breaks = seq(0, 30, 1),
                     limits = c(0, 30),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase_excel.tif", 
       plot = phase_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_4_Article_phase_excel.tif"))
```

```{r fig4-phase-greys}
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(Phase,n), y = n)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  xlab("") +
  geom_text(aes(y = 10, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 30, 5),
                     minor_breaks = seq(0, 30, 1),
                     limits = c(0, 30),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase_greys.tif", 
       plot = phase_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_4_Article_phase_greys.tif"))
```

### Three phase

```{r fig4-phase-3-blues}
phase_3_bar <- summary_phase_condensed %>% 
  ggplot(aes(x = reorder(Phase_condensed, n), y = n)) +
  geom_bar(stat = "identity", fill = blues[4], colour = "black") +
  xlab("") +
  geom_text(aes(y = 10, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 60, 5),
                     minor_breaks = seq(0, 60, 1),
                     limits = c(0, 60),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase_3_blues.tif", 
       plot = phase_3_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_4_Article_phase_3_blues.tif"))
```

```{r fig4-phase-3-excel}
phase_3_bar <- summary_phase_condensed %>% 
  ggplot(aes(x = reorder(Phase_condensed, n), y = n)) +
  geom_bar(stat = "identity", fill = excel_pal[3], colour = "black") +
  xlab("") +
  geom_text(aes(y = 10, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 60, 5),
                     minor_breaks = seq(0, 60, 1),
                     limits = c(0, 60),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase_3_excel.tif", 
       plot = phase_3_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_4_Article_phase_3_excel.tif"))
```

```{r fig4-phase-3-greys}
phase_3_bar <- summary_phase_condensed %>% 
  ggplot(aes(x = reorder(Phase_condensed, n), y = n)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  xlab("") +
  geom_text(aes(y = 10, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 60, 5),
                     minor_breaks = seq(0, 60, 1),
                     limits = c(0, 60),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure 4. TOP" ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_4_Article_phase_3_greys.tif", 
       plot = phase_3_bar, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_4_Article_phase_3_greys.tif"))
```

## Figure 5 Articles by Source-Year

Done with bars side-by-side or stacked.

### besides

```{r source_year_prep}
summary_tissue <- dat %>% 
  group_by(Source, Year = factor(Year), 
           .drop = FALSE, .preserve = TRUE) %>% 
  summarise(n = length(Source))

```

```{r fig5-source-year-blues}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", colour = "black", 
           position = position_dodge()) +
  scale_fill_manual(values = blues[2:7], 
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 12, 1),
                    minor_breaks = seq(0, 12, 1),
                    limits = c(0, 12),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year_blues.tif", 
       plot = source_year, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_5_Article_source_year_blues.tif"))
```

```{r fig5-source-year-excel}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", colour = "black", 
           position = position_dodge()) +
  scale_fill_manual(values = excel_pal, 
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 12, 1),
                    minor_breaks = seq(0, 12, 1),
                    limits = c(0, 12),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year_excel.tif", 
       plot = source_year, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_5_Article_source_year_excel.tif"))
```

```{r fig5-source-year-greys}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", colour = "black", 
           position = position_dodge()) +
  scale_fill_manual(values = greys, 
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 12, 1),
                    minor_breaks = seq(0, 12, 1),
                    limits = c(0, 12),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year_greys.tif", 
       plot = source_year, 
       device = device,
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_5_Article_source_year_greys.tif"))
```

### stacked
```{r fig5-source-year-stack-blues}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = blues[2:7], 
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 12, 1),
                    minor_breaks = seq(0, 12, 1),
                    limits = c(0, 12),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year_stack_blues.tif", 
       plot = source_year, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_5_Article_source_year_stack_blues.tif"))
```

```{r fig5- source-year-stack-excel}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = excel_pal, 
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 12, 1),
                    minor_breaks = seq(0, 12, 1),
                    limits = c(0, 12),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year_stack_excel.tif", 
       plot = source_year, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_5_Article_source_year_stack_excel.tif"))
```

```{r fig5-source-year-stack-greys}
source_year <- summary_tissue %>% 
  ggplot(aes(x = Year, y = n, fill = Source)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = greys, 
                    guide = guide_legend(nrow = 2, title = NULL )) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 12, 1),
                    minor_breaks = seq(0, 12, 1),
                    limits = c(0, 12),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Year") +
  labs(title = "Wilson, Alison, Figure 5. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.position = "top",
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_5_Article_source_year_stack_greys.tif", 
       plot = source_year, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
grid.raster(readTIFF("figs/fig_5_Article_source_year_stack_greys.tif"))
```

## Figure 6 Articles by Test quality

```{r test-quality-prep}
summary_test <- dat %>% 
  group_by(test_results) %>% 
  summarise(n = length(test_results))

summary_test$x <- 1
tot <- sum(summary_test$n)
n <- summary_test$n
summary_test$lab_y <- c(n[1]/2,
                        n[1] + n[2]/2,
                        n[1] + n[2] + n[3]/2,
                        n[1] + n[2] + n[3] + n[4]/2,
                        n[1] + n[2] + n[3] + n[4] + n[5]/2)
```

```{r fig6-test-quality}
test_quality <- summary_test %>% 
  ggplot(aes(x = x, y = n, fill = rev(test_results)) ) +
  geom_bar(stat = "identity", 
           colour = "black", 
           width = 2) +
  geom_text(aes(x = x, y = lab_y, label = n),
            size = 6) +
  geom_text(aes(x = 2.2, y = lab_y, label = test_results),
            hjust = 0, size = 6 ) +
  scale_fill_manual(values = c(greys[3], greys[5], greys[6], greens[2], greens[4]),
                    name = "Test results") +
  xlim(0, 12)  +
  labs(title = "Wilson, Alison, Figure 7. TOP" ) +
  theme(panel.background = element_blank(),
        legend.position = "none",
        axis.title = element_blank(),
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        panel.grid.minor.y = element_line(colour = "white", size = 0.3),
        panel.grid.major.x = element_blank()) 
ggsave("figs/fig_6_Article_test_quality.tif", 
       plot = test_quality, 
       device = device,
       width = fig_w_1.5, 
       height = fig_h,
       units = units,
       dpi = dpi)

```

## Figure 6 Articles by Phase and Test quality reported

### Stacked by number of articles
```{r}

summary_phase_test <- dat %>%
  group_by(Phase_condensed, test_results) %>% 
  summarise(n = length(Phase))
```

```{r fig7-phase-test-n}
phase_test <- summary_phase_test %>% 
  ggplot(aes(x = Phase_condensed, y = n, fill = test_results)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = c(greens[4], greens[2], greys[6], greys[5],greys[3]),
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 55, 10),
                     minor_breaks = seq(0, 55, 5),
                     limits = c(0, 55),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
  labs(title = "Wilson, Alison, Figure 7. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.title = element_text(size = 8, colour = "black"),
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.2, "in"),
        legend.position = c(0.7, 0.8))
ggsave("figs/fig_7_Article_phase_test_quality_by_N.tif", 
       plot = phase_test, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

### Stacked by percent of article

```{r fig7-phase-test-pc}
phase_test2 <- summary_phase_test %>% 
  ggplot(aes(x = Phase_condensed, y = n)) +
  geom_bar(aes(fill = test_results), 
           stat = "identity", 
           colour = "black", 
           position = "fill") +
  scale_fill_manual(values = c(greens[4], greens[2]
                               , greys[6], greys[5],greys[3]),
                    name = "Test results", 
                    guide = guide_legend(nrow = 3, title = NULL )) +
  scale_y_continuous(name = "Proportion of Articles by Phase",
                     breaks = seq(0, 1.2, 0.2),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
  geom_text(data = summary_phase_condensed,
            aes(x = Phase_condensed, y = 1, vjust = -0.4,
                label = paste("n = ",n))) +
  labs(title = "Wilson, Alison, Figure 7. TOP" ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        legend.title = element_text(size = 6, colour = "black"),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.2, "in"),
        legend.position = "top") +
  coord_cartesian(clip = 'off')
ggsave("figs/fig_7_Article_phase_test_quality_by_PC.tif", 
       plot = phase_test2, 
       device = device,
       width = fig_w, 
       height = fig_h,
       units = units,
       dpi = dpi)
```

## Figure 9

There are some inconsistencies in the data file columns for markers (most have three associated columns (value, status and an unnamed one which appears to be a qualifier for the value and which I have named qualifier) but some have one a value but not status. It's not obv what the status should be in some cases.

The markers in this figure thus differ because I'm missed out the unknown status ones.

There's also a single "not stated" value remianing

```{r}
dat2 <- dat %>% 
  select(contains("_status")) %>% 
  rename_all(funs(stringr::str_replace(., fixed("_status"),"") )) %>% 
  gather(key = "marker", value = "status") %>%
  mutate(ISCT_pheno = recode_factor(marker,
                                CD73 = TRUE,
                                CD90 = TRUE,
                                CD105 = TRUE,
                                CD34 = TRUE,
                                CD45 = TRUE,
                                CD11b = TRUE,
                                CD14 = TRUE,
                                CD79a = TRUE,
                                CD19 = TRUE,
                                `HLA-DR` = TRUE,
                                .default = FALSE))  
# relevel the factor
dat2$marker <- fct_relevel(dat2$marker,
                           "viability", 
                           "CD73", 
                           "CD90", 
                           "CD105",
                           "CD34", 
                           "CD45", 
                           "CD11b",
                           "CD14", 
                           "CD79a",
                           "CD19",
                           "HLA-DR",
                           "CD133",
                           "CD166",
                           "CD29",
                           "CD44")
charc_marker <- dat2 %>%
  group_by(marker, status) %>% 
  summarise(n = length(marker))
charc_marker$marker <- fct_relevel(charc_marker$marker,
                           "viability", 
                           "CD73", 
                           "CD90", 
                           "CD105",
                           "CD34", 
                           "CD45", 
                           "CD11b",
                           "CD14", 
                           "CD79a",
                           "CD19",
                           "HLA-DR",
                           "CD133",
                           "CD166",
                           "CD29",
                           "CD44")
```

```{r}
charc_marker %>% 
  ggplot(aes(x = marker, y = n)) +
  geom_bar(aes(fill = status), 
           stat = "identity",
           position = "fill",
           colour = "white") +
  scale_fill_manual(values = blues[4:7], name = "Status") +
  scale_y_continuous(name = "Proportion of Articles") +
  scale_x_discrete(expand = c(0, 0), name = "Marker") +
  theme(panel.background = element_rect(fill = blues[2]),
        panel.grid = element_blank(),)

```

## Relationship between MOA and stem/stromal

```{r}
moa_isct_summary <- dat %>% 
  group_by(MOA, stem_stromal) %>% 
  summarise(n = length(MOA))

res <- table(dat$stem_stromal, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)
test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))

```

```{r moa-isct-blues}
moa_isct <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = blues[2:5], 
                    guide = guide_legend(nrow = 4, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x. TOP",
       caption = test_anno) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_x_moa_isct_blues.tif", 
       plot = moa_isct, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r moa-isct-excel}
moa_isct <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = excel_pal[1:4], 
                    guide = guide_legend(nrow = 4, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x. TOP",
       caption = test_anno ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in")) 
 ggsave("figs/fig_x_moa_isct_excel.tif", 
       plot = moa_isct, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r moa-isct-greys}
moa_isct <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = greys[3:6], 
                    guide = guide_legend(nrow = 4, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x. TOP",
       caption = test_anno ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_x_moa_isct_greys.tif", 
       plot = moa_isct, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

## Relationship between types of differentiation (OAC) and MOA

```{r}
table_info <- "There are five combinations for O, A and C potency. 'y' is demonstrated, 'm' is claimed but not demonstrated, 'n' is not mentioned. These were recoded into a single variable, 'potency'."

dat %>% 
  group_by(O, A, C, potency) %>% 
  summarise(n = length(O)) %>% 
  kable(caption = table_info) %>% 
  kable_styling() %>% 
  save_kable("figs/recode_differentiation_capacity.png")

```

```{r}
res <- table(dat$potency, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)

test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))
```

```{r moa-potency-prep}
moa_potency_summary <- dat %>% 
  group_by(potency, MOA) %>% 
  summarise(n = length(MOA))
```

```{r moa_potency-blues}
moa_potency <- moa_potency_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = potency)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = blues[2:6], 
                    guide = guide_legend(nrow = 5, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x2. TOP",
       caption = test_anno) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_x2_moa_potency_blues.tif", 
       plot = moa_potency, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
```

```{r moa_potency-excel}
moa_potency <- moa_potency_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = potency)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = excel_pal[2:6], 
                    guide = guide_legend(nrow = 5, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x2. TOP",
       caption = test_anno) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_x2_moa_potency_excel.tif", 
       plot = moa_potency, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
```

```{r moa_potency-greys}
moa_potency <- moa_potency_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = potency)) +
  geom_bar(stat = "identity", colour = "black") +
  scale_fill_manual(values = greys[2:6], 
                    guide = guide_legend(nrow = 5, title = NULL)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 45, 5),
                     minor_breaks = seq(0, 45, 5),
                     limits = c(0, 45),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Mechanism of action") +
  labs(title = "Wilson, Alison, Figure x2. TOP",
       caption = test_anno) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4),
        plot.caption = element_text(size = 4, hjust = 0),
        legend.position = c(0.7, 0.8),
        legend.text = element_text(size = 6, colour = "black"),
        legend.key.size = unit(0.1, "in"))
ggsave("figs/fig_x2_moa_potency_greys.tif", 
       plot = moa_potency, 
       device = device,
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)
```

## Claimed compliance vs categ compliance

```{r}
vals <- unique(dat$`ISCT compliance`)
ref <- dat$`Ref #`[is.na(dat$`ISCT compliance`)]
```

The values in `ISCT compliance` are: `r vals`
Probably the empty cell (formal NA in r) should be  "n/a" (a string). It's reference
`r ref`

```{r}

dat %>% 
  group_by(`ISCT compliance`,`ISCT claimed`) %>%
  summarise(n = length(`ISCT claimed`)) %>% 
  kable() %>% 
  kable_styling()

```
These results are so clear (all 14 of the papers that claim compliance do not demonstrate it; no other papers demonstrate it) that no further analysis is necessary.

## Relevance for indication, stem vs stromal
There are 50 different indications mostly with a single ref. This makes it difficult to deterimine if there is any relationalship between the nature of the indication and whether treatment was with stem of stromal cells.

```{r}
dat %>% 
  group_by(Indication, `Stem/Stromal`) %>% 
  summarise(n = length(`Stem/Stromal`)) %>% 
  ggplot(aes(x = reorder(Indication, n), y = n, fill = `Stem/Stromal`)) +
  geom_col() +
  coord_flip()
```

```{r}
ind <- dat %>% 
  group_by(Indication) %>% 
  summarise(n = length(potency)) %>% 
  filter(n > 1)

dat %>% 
  group_by(Indication,potency) %>% 
  summarise(n = length(Indication)) %>% 
  filter(n > 1)
```