### 2020-06-09 paper2.Rmd
transfer

---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r pkg}
library(tidyverse)
library(waffle)
library(RColorBrewer)
library(readxl)
library(kableExtra)
```

```{r}
blues <- brewer.pal(9,"Blues")
greens <- brewer.pal(9,"Greens")
excel_pal <- c("#CEDBE6", "#3494BA", "#58B6C0", "#75BDA7", "#84ACB6", "#2683C6", "#373545")
```

```{r import}
file <- here::here("data", "Meta-Analysis Data Collection.xlsx")

# added a row for column names more useful
# # skipped first two rows of of old col names

dat <- read_excel(file, sheet = "Data")

```

Note: there is a 85th line containing only a note in the `NOTES` variable - presumably it belongs elsewhere and was excluded. The note was "compares different papers on bone marrow MSCs, doesn't say anything about prep of cells"

```{r tidy-data}
# recode country so those with one paper are
# "Other"
dat <- dat %>%
  mutate(Country2 = recode_factor(Country,
                                Brazil = "Brazil",
                                China = "China",
                                Denmark = "Denmark",
                                Egypt = "Egypt",
                                India = "India",
                                Japan = "Japan",
                                Korea = "Korea",
                                Netherlands = "Netherlands",
                                Poland = "Poland",
                                `Rep Korea` = "Rep Korea",
                                Russia = "Russia",
                                Spain = "Spain",
                                Taiwan = "Taiwan",
                                USA = "USA",
                                .default = "Other"))

dat <- dat %>%
  mutate(Phase_condensed = recode_factor(Phase,
                               I = "I",
                               `I/IIa` = "I",
                                II = "II",
                                IIa = "II",
                               IIb = "II",
                               III = "III"))

# test results - there is no "t" in the data file and there are more
# blanks values than t's in draft paper fig 6
dat <- dat %>%
  mutate(test_results = recode_factor(`Test Results?`,
                                       n = "2. Average value",
                                       s = "4. Stated as 'standard phenotype'",
                                       o = "5. Not discussed",
                                       y = "1. Phenotypic markers reported",
                                       t = "3. Tests done but not reported"))
dat$test_results <- fct_relevel(dat$test_results,
                              "1. Phenotypic markers reported", 
                              "2. Average value",
                              "3. Tests done but not reported",
                              "4. Stated as 'standard phenotype'",
                              "5. Not discussed")

# dat <- dat %>%
#   mutate(Test.Results.2 = recode_factor(Test.Results.,
#                                        `Average value` = "Bulk values",
#                                        `Phenotypic markers reported` = "Phenotypic markers reported",
#                                        .default = "No characterisation data"))

######### code for "Data Tab copy v1.csv"
# # column indices for markers
# firstmarker <- which(colnames(dat) == "CD73")
# lastmarker <- which(colnames(dat) == "SSEA.4")
# # vector of marker names
# markers <- names(dat[firstmarker:lastmarker])
# 
# datnum <- dat %>% 
#   select(markers) %>% 
#   mutate_all(as.numeric) 
# 
# # add the non marker columns
# notmarkers <- which(!(colnames(dat) %in% markers))
# datnum <- cbind(dat[notmarkers], datnum)
# 
# 
# dat$Stem.Stromal <- factor(dat$Stem.Stromal)
# dat$Year <- factor(dat$Year)
# datnum$Stem.Stromal <- factor(datnum$Stem.Stromal)
# datnum$Year <- factor(datnum$Year)

```
# Some figures

Resolutions, colours, font types and size, labels, background colours and gridlines etc can be removed / adapted easily.

Once we've decided on figures, I'll turn them in to high res and big fonts suitable for shrinking.

## fig 1 - search terms

Funnel plot - could be adapted to include the reviews adjacent and aligned with the articles plot.

> to do: Add the Reviews diagram

```{r data-prep-fig1}
# create data
# short names for proof of method
orig <- data.frame(step_names = c("1.mesenchymal",
                                  "2.article",
                                  "3.cell tissue etc",
                                  "4.clinical etc",
                                  "5.not biotech etc",
                                  "6.exc conf etc"),
                   n = c(65706,
                         47031,
                         7814,
                         5502,
                         2055,
                         1986))

orig$step_names <- fct_relevel(orig$step_names,
                               "mesenchymal",
                               "article",
                               "cell tissue etc",
                               "clinical etc",
                               "not biotech etc",
                               "exc conf etc")

# prep polygons that connect
to_poly_orig <- orig %>%
  arrange(n) %>%
  mutate(
    group = row_number(),
    x1 = 0 + ((1 - n)/2),
    x4 = n + ((1 - n)/2),
    x2 = lag(x1), 
    x3 = lag(x4)) %>%
  mutate(
    x2 = ifelse(is.na(x2), x1, x2),
    x3 = ifelse(is.na(x3), x4, x3)
  ) %>%
  gather(xpoint, x, -step_names,-group, -n) %>%
  arrange(group, xpoint) %>%
  mutate(y = ifelse(xpoint %in% c("x2", "x3"), group, group + 1))

# EXAMPLE
# Create the placement of the stages in the plot,
#  by just adding a 0.5 to each row number

labels <- orig %>%
  arrange(n) %>%
  mutate(y = row_number() + 0.5)

# modify the labels
labels$step_names2 <- c( "EXCLUDE conference proceedings,\nretracted papers",
                         "EXCLUDE CATEGORIES (NOT\nbiotechnology applied microbiology oncology\nhaematology, engineering biomedical)",
                         "REFINE BY SOURCE TITLES\n(EXCLUDE NOT clinical/\ntranslation/regenerative medicine)",
                         "REFINE BY CATEGORY\ncell tissue engineering\nand transplantation",
                         "REFINE BY DOCUMENT TYPE\narticle",
                         "mesenchymal stem cells OR\nmesenchymal stromal cells OR\nMSC AND characterisation AND clinical")
```

```{r fig1}
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               color = "grey10") +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 8,
            label.size = 0,
            label.padding = unit(0.2, "lines"),
            label.r = unit(0.2, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 20, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
    scale_fill_manual(values = blues)

ggsave("figs/fig_1_Article_search_funnel_blues.png", 
       plot = article_funnel, 
       device = "png",
       width = 30, 
       height = 20,
       units = "cm")
```

## fig 3 papers by country

### bar
Better, With all the countries. Horizontal makes it easier to read Country names without rotating text.

```{r}
fig_3_Articles_by_country_blues <- summary_country %>% 
  ggplot(aes(x = reorder(Country, n), y = n)) +
  geom_bar(stat = "identity", fill = blues[4], color = "black") +
  xlab("") +
  scale_y_continuous(name = "Number of articles", 
                    breaks = seq(0, 18, 1),
                    minor_breaks = seq(0, 18, 1),
                    limits = c(0, 18.5),
                    expand = c(0, 0)) +
  scale_x_discrete() +
  theme_classic() +
  theme(axis.ticks.y = element_blank(),
        axis.text  = element_text(size = 20, colour = "black"),
        axis.title.x = element_text(size = 22)) +
  coord_flip()
ggsave("figs/fig_3_Articles_by_country_blues.png", 
       plot = fig_3_Articles_by_country_blues, 
       device = "png",
       width = 25, 
       height = 25,
       units = "cm") 
```

## fig 4 paper by phase

```{r waffle-phase-prep}
summary_phase <- dat %>%
  group_by(Phase) %>% 
  summarise(n = length(Phase),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
summary_phase <- summary_phase %>%
  mutate(Phase_condensed = recode_factor(Phase,
                               I = "I",
                               `I/IIa` = "I",
                                II = "II",
                                IIa = "II",
                               IIb = "II",
                               III = "III"))

summary_phase_condensed <- dat %>%
  group_by(Phase_condensed) %>% 
  summarise(n = length(Phase_condensed),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

```

Phase can comprise six or three categoires: `r unique(summary_phase$Phase)`; or three categories: `r unique(summary_phase_condensed$Phase_condensed)`

## waffle

### Six phase

```{r}
parts <- summary_phase$n
names(parts) <- summary_phase$Phase
waffle(parts,
       rows = 7,
       colors = blues[4:9])

```

### Three phase

```{r}
parts <- summary_phase_condensed$n
names(parts) <- summary_phase_condensed$Phase_condensed
waffle(parts,
       rows = 7,
       colors = blues[c(4, 6, 8)])

```

## Bar
as a bar. can add percentages as annotation if desired. can order bars as required.

### Six phase

```{r}
summary_phase %>% 
  ggplot(aes(x = reorder(Phase,n), y = n)) +
  geom_bar(stat = "identity", fill = blues[4]) +
  xlab("") +
  geom_text(aes(y = n / 2, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 30, 5),
                    minor_breaks = seq(0, 30, 1),
                    limits = c(0, 30),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
    theme(panel.background = element_rect(fill = blues[2]),
        panel.grid.minor.x = element_line(colour = "white", size = 0.3),
        panel.grid.major.y = element_blank(),
        axis.ticks = element_blank()) +
  coord_flip()
```

### Three phase

```{r}
summary_phase_condensed %>% 
  ggplot(aes(x = reorder(Phase_condensed, n), y = n)) +
  geom_bar(stat = "identity", fill = blues[4]) +
  xlab("") +
  geom_text(aes(y = n / 2, label = pc) ) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 60, 10),
                    minor_breaks = seq(0, 60, 5),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
    theme(panel.background = element_rect(fill = blues[2]),
        panel.grid.minor.x = element_line(colour = "white", size = 0.3),
        panel.grid.major.y = element_blank(),
        axis.ticks = element_blank()) +
  coord_flip()
```

(will move percent if this figure is selected)

### Three and six phase both indicated
```{r}
summary_phase %>% 
  ggplot(aes(x = reorder(Phase_condensed, n), y = n, fill = Phase)) +
  geom_bar(stat = "identity", colour = "white") +
  xlab("") +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 60, 10),
                    minor_breaks = seq(0, 60, 5),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_manual(values = blues[4:9]) +
    theme(panel.background = element_rect(fill = blues[2]),
        panel.grid.minor.x = element_line(colour = "white", size = 0.3),
        panel.grid.major.y = element_blank(),
        axis.ticks = element_blank()) +
  coord_flip()
```

## fig 5 paper msc source per year

```{r}
summary_tissue <- dat %>% 
  group_by(Year, Source) %>% 
  summarise(n = length(Source))

```

```{r}
summary_tissue %>% 
  ggplot(aes(x = factor(Year), y = n, fill = Source)) +
  geom_bar(stat = "identity", colour = "white") +
  scale_fill_manual(values = blues[4:9]) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 15, 1),
                    minor_breaks = seq(0, 15, 1),
                    limits = c(0, 15),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "") +
    theme(panel.background = element_rect(fill = blues[2]),
        panel.grid.minor.y = element_line(colour = "white", size = 0.3),
        panel.grid.major.x = element_blank())
```

## fig 6

```{r}
summary_test <- dat %>% 
  group_by(test_results) %>% 
  summarise(n = length(test_results))

summary_test$x <- 1
tot <- sum(summary_test$n)
n <- summary_test$n
summary_test$lab_y <- c(n[1]/2,
                        n[1] + n[2]/2,
                        n[1] + n[2] + n[3]/2,
                        n[1] + n[2] + n[3] + n[4]/2,
                        n[1] + n[2] + n[3] + n[4] + n[5]/2)
```

Might you want test colours that indicate 'good' and 'bad' (green and grey maybe)
```{r}
summary_test %>% 
  ggplot(aes(x = x, y = n, fill = rev(test_results)) ) +
  geom_bar(stat = "identity", colour = "white", width = 0.5) +
  geom_text(aes(x = x, y = lab_y, label = n), size = 4) +
  geom_text(aes(x = 1.3, y = lab_y, label = test_results), hjust = 0 ) +
  scale_fill_manual(values = blues[3:7], name = "Test results") +
  scale_x_discrete(breaks = seq(0, 3, 1),
                   limits = c(0, 3)) +
    theme(panel.background = element_blank(),
          legend.position = "none",
          axis.title = element_blank(),
          axis.text = element_blank(), 
          axis.ticks = element_blank(),
          panel.grid.minor.y = element_line(colour = "white", size = 0.3),
          panel.grid.major.x = element_blank()) 
```

## figire 7 new phase and tests reported

### stacked by numbers
```{r}

summary_phase_test <- dat %>%
  group_by(Phase_condensed, test_results) %>% 
  summarise(n = length(Phase))
```
Might you want test colours that indicate 'good' and 'bad' (green and grey maybe)

```{r}
summary_phase_test %>% 
  ggplot(aes(x = Phase_condensed, y = n, fill = test_results)) +
  geom_bar(stat = "identity", colour = "white") +
  scale_fill_manual(values = blues[3:7], name = "Test results") +
  scale_y_continuous(name = "Number of Articles",
                    breaks = seq(0, 55, 10),
                    minor_breaks = seq(0, 55, 5),
                    limits = c(0, 55),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
    theme(panel.background = element_rect(fill = blues[2]),
        panel.grid.minor.y = element_line(colour = "white", size = 0.3),
        panel.grid.major.x = element_blank())
```

### stacked by percent

```{r}
summary_phase_test %>% 
  ggplot(aes(x = Phase_condensed, y = n)) +
  geom_bar(aes(fill = test_results), 
           stat = "identity", 
           colour = "white", 
           position = "fill") +
  scale_fill_manual(values = blues[3:7], name = "Test results") +
  scale_y_continuous(name = "Proportion of Articles by Phase",
                     breaks = seq(0, 1.1, 0.1),
                     # minor_breaks = seq(0, 1.1, 0.1),
                     # limits = c(0, 1.2),
                     expand = c(0, 0.1)) +
   scale_x_discrete(expand = c(0, 0), name = "Phase") +
    theme(panel.background = element_rect(fill = blues[2]),
                panel.grid = element_blank()) +
  geom_text(data = summary_phase_condensed, 
            aes(x = Phase_condensed, y = 1.05,
                label = paste("n = ",n)))
```

## Figure 9

There are some inconsistencies in the data file columns for markers (most have three associated columns (value, status and an unnamed one which appears to be a qualifier for the value and which I have named qualifier) but some have one a value but not status. It's not obv what the status should be in some cases.

The markers in this figure thus differ because I'm missed out the unknown status ones.

There's also a single "not stated" value remianing

```{r}
dat2 <- dat %>% 
  select(contains("_status")) %>% 
  rename_all(funs(stringr::str_replace(., fixed("_status"),"") )) %>% 
  gather(key = "marker", value = "status") %>%
  mutate(ISCT_pheno = recode_factor(marker,
                                CD73 = TRUE,
                                CD90 = TRUE,
                                CD105 = TRUE,
                                CD34 = TRUE,
                                CD45 = TRUE,
                                CD11b = TRUE,
                                CD14 = TRUE,
                                CD79a = TRUE,
                                CD19 = TRUE,
                                `HLA-DR` = TRUE,
                                .default = FALSE))  
# relevel the factor
dat2$marker <- fct_relevel(dat2$marker,
                           "viability", 
                           "CD73", 
                           "CD90", 
                           "CD105",
                           "CD34", 
                           "CD45", 
                           "CD11b",
                           "CD14", 
                           "CD79a",
                           "CD19",
                           "HLA-DR",
                           "CD133",
                           "CD166",
                           "CD29",
                           "CD44")
charc_marker <- dat2 %>%
  group_by(marker, status) %>% 
  summarise(n = length(marker))
charc_marker$marker <- fct_relevel(charc_marker$marker,
                           "viability", 
                           "CD73", 
                           "CD90", 
                           "CD105",
                           "CD34", 
                           "CD45", 
                           "CD11b",
                           "CD14", 
                           "CD79a",
                           "CD19",
                           "HLA-DR",
                           "CD133",
                           "CD166",
                           "CD29",
                           "CD44")
```

```{r}
charc_marker %>% 
  ggplot(aes(x = marker, y = n)) +
  geom_bar(aes(fill = status), 
           stat = "identity",
           position = "fill",
           colour = "white") +
  scale_fill_manual(values = blues[4:7], name = "Status") +
  scale_y_continuous(name = "Proportion of Articles") +
  scale_x_discrete(expand = c(0, 0), name = "Marker") +
  theme(panel.background = element_rect(fill = blues[2]),
        panel.grid = element_blank(),)

```

## Relationship between MOA and stem/stromal

```{r}
moa_isct_summary <- dat %>% 
  group_by(MOA, `Stem/Stromal`) %>% 
  summarise(n = length(MOA))

res <- table(dat$`Stem/Stromal`, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)

```

There is no significant association between MOA and cell categorisation (`r res$method`:  $p$ = `r res$p.value`).

```{r}
moa_isct_summary %>% 
  ggplot(aes(x = factor(MOA), y = n, fill = `Stem/Stromal`)) +
  geom_bar(stat = "identity", colour = "white") +
  scale_fill_manual(values = blues[5:8]) +
  scale_y_continuous(name = "Number of Articles", 
                    breaks = seq(0, 45, 5),
                    minor_breaks = seq(0, 45, 5),
                    limits = c(0, 45),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "MOA") +
    theme(panel.background = element_rect(fill = blues[2]),
        panel.grid.minor.y = element_line(colour = "white", size = 0.3),
        panel.grid.major.x = element_blank())
```

## Relationship between types of differentiation (OAC) and MOA

There are five combinations for O, A and C potency. `y` is demonstrated, `m` is claimed but not demonstrated, `n` is not mentioned. These were recoded into a single variable, `potency` as follows:
```{r}
# there are five combinations of 
dat <- dat %>%
  mutate(potency = ifelse(O == "y" &
                            A == "y" &
                            C == "y",
                          "demonstrated tripotency",
                          ifelse(O == "n" &
                                   A == "n" &
                                   C == "n",
                                 "not mentioned",
                                 ifelse(O == "y" &
                                          A == "y" &
                                          C == "n",
                                        "demonstrated bipotency (OA)",
                                        ifelse(O == "m"
                                               & A == "m" &
                                                 C == "n",
                                               "claimed bipotency (OA)",
                                               ifelse(O == "m" &
                                                  A == "m" &
                                                  C == "m",
                                                  "claimed tripotency", NA))))))
```

```{r}
dat %>% 
  group_by(O, A, C, potency) %>% 
  summarise(n = length(O)) %>% 
  kable(format = "html") %>% 
  kable_styling()

```

```{r}
res <- table(dat$potency, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)

```

There is no significant association between MOA and cell categorisation (`r res$method`:  $p$ = `r res$p.value`).

```{r}
moa_potency_summary <- dat %>% 
  group_by(potency, MOA) %>% 
  summarise(n = length(MOA))
```

```{r}
moa_potency_summary %>% 
  ggplot(aes(x = factor(MOA), y = n, fill = potency)) +
  geom_bar(stat = "identity", colour = "white") +
  scale_fill_manual(values = blues[4:8]) +
  scale_y_continuous(name = "Number of Articles",
                    breaks = seq(0, 45, 5),
                    minor_breaks = seq(0, 45, 5),
                    limits = c(0, 45),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "MOA") +
    theme(panel.background = element_rect(fill = blues[2]),
        panel.grid.minor.y = element_line(colour = "white", size = 0.3),
        panel.grid.major.x = element_blank())
```

## Claimed compliance vs categ compliance

```{r}
vals <- unique(dat$`ISCT compliance`)
ref <- dat$`Ref #`[is.na(dat$`ISCT compliance`)]
```

The values in `ISCT compliance` are: `r vals`
Probably the empty cell (formal NA in r) should be  "n/a" (a string). It's reference
`r ref`

```{r}

dat %>% 
  group_by(`ISCT compliance`,`ISCT claimed`) %>%
  summarise(n = length(`ISCT claimed`)) %>% 
  kable() %>% 
  kable_styling()

```
These results are so clear (all 14 of the papers that claim compliance do not demonstrate it; no other papers demonstrate it) that no further analysis is necessary.

## Relevance for indication, stem vs stromal
There are 50 different indications mostly with a single ref. This makes it difficult to deterimine if there is any relationalship between the nature of the ndication and whether treatment was with stem of stromal cells.

```{r}
dat %>% 
  group_by(Indication, `Stem/Stromal`) %>% 
  summarise(n = length(`Stem/Stromal`)) %>% 
  ggplot(aes(x = reorder(Indication, n), y = n, fill = `Stem/Stromal`)) +
  geom_col() +
  coord_flip()
```

```{r}
ind <- dat %>% 
  group_by(Indication) %>% 
  summarise(n = length(potency)) %>% 
  filter(n > 1)

dat %>% 
  group_by(Indication,potency) %>% 
  summarise(n = length(Indication)) %>% 
  filter(n > 1)
```