### 2020-08-10 paper2_figures.Rmd
fig updates for data update

---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  wmf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

# Figures specification
Vector-based files 
150 to 300 dots per inch (dpi) for initial submission. A higher resolution applies for figures submitted at the revision stage.

Fit on single 8.5" × 11" or A4 paper
Preferably at:
1 column - 3.5 inches
1.5 columns - 5.0 inches
2 columns = 7.3 inches

The figure title should be at the beginning of the figure legend
Include the fig id on same page that includes the figure

```{r pkg}
library(tidyverse)
# library(readxl)
# library(kableExtra)
library(grid)
library(patchwork)
library(janitor)

```

```{r}

# palette
pal <- viridisLite::viridis(6)

# figure saving settings
units = "in"
fig_w <- 3.5 # one col
fig_w_1.5 <- 5 # 1.5 col
fig_w_2 <- 7.3 # 2 col
fig_h <- fig_w

```

## write to file
```{r}
# file <- "data/checked_processed_aw_corrections.csv"
# write_csv(dat, file)
```

# Read in the processed and corrected data

```{r}
file <- "data/checked_processed_aw_corrections.csv"
dat <- read_csv(file, skip = 1)
# write(names(dat), "data/varnames.txt")

```

```{r}
# put levels in
source("levels.R")
```

# Figures

## Fig 1 - Search

### prep

```{r data-prep-fig1}
# create data for funnel 1A
# short names for proof of method
orig <- data.frame(step_names = c("1.mesenchymal",
                                  "2.article",
                                  "3.cell tissue etc",
                                  "4.clinical etc",
                                  "5.not biotech etc",
                                  "6.exc conf etc"),
                   n = c(65706,
                         47031,
                         7814,
                         5502,
                         2055,
                         1986))

# prep polygons that connect
to_poly_orig <- orig %>%
  arrange(n) %>%
  mutate(
    group = row_number(),
    x1 = 0 + ((1 - n)/2),
    x4 = n + ((1 - n)/2),
    x2 = lag(x1), 
    x3 = lag(x4)) %>%
  mutate(
    x2 = ifelse(is.na(x2), x1, x2),
    x3 = ifelse(is.na(x3), x4, x3)
  ) %>%
  gather(xpoint, x, -step_names,-group, -n) %>%
  arrange(group, xpoint) %>%
  mutate(y = ifelse(xpoint %in% c("x2", "x3"), group, group + 1))

# EXAMPLE
# Create the placement of the stages in the plot,
#  by just adding a 0.5 to each row number

labels <- orig %>%
  arrange(n) %>%
  mutate(y = row_number() + 0.5)

# modify the labels
labels$step_names2 <- c( "EXCLUDE,\nconference proceedings,\nretracted papers",
                         "EXCLUDE CATEGORIES (NOT\nbiotechnology, applied\nmicrobiology, oncology\nhaematology,\nengineering biomedical)",
                         "REFINE BY SOURCE TITLES\n(EXCLUDE NOT clinical/\ntranslation/regenerative,\nmedicine)",
                         "REFINE BY CATEGORY\ncell tissue engineering\nand transplantation",
                         "REFINE BY DOCUMENT\nTYPE article",
                         "mesenchymal stem cells OR\nmesenchymal stromal cells\nOR MSC AND characterisation,\nAND clinical")

# read data for 1B
fig1b <- read.table("data/fig1b.txt", header = TRUE)

```

### each panel of fig 1
```{r fig1}
# figure 1 a
article_funnel <- to_poly_orig %>%
  ggplot() +
  geom_polygon(aes(x,
                   y,
                   group = group,
                   fill = step_names),
               size = 0) +
  geom_label(aes(x = 0,
                y = y,
                label = n),
            data = labels, 
            size = 3,
            label.size = 0,
            label.padding = unit(0.1, "lines"),
            label.r = unit(0.1, "lines")) +
  scale_y_continuous(breaks = labels$y,
                     labels = labels$step_names2) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y  = element_text(size = 7, colour = "black", hjust = 0, face = "plain"),
        panel.spacing = unit(0, "lines"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
    scale_fill_manual(values = rev(pal))

# fig 1b

data_collection <- fig1b %>% 
  ggplot() +
  geom_text(aes(x = x, y = y, label = item),
            size = 2.5,nudge_y = 0.1,
            hjust = "left") +
  facet_grid(. ~ data_element) +
  scale_x_continuous(limits = c(0, 7)) +
  theme_minimal() +
  theme(legend.position = "none",
        strip.text = element_text(hjust = 0, size = 5.5, colour = "white", face = "bold"),
        strip.background = element_rect(fill = pal[1], colour = pal[1], size = 0.05),
        panel.border = element_rect(colour = "black", fill = NA),
        panel.spacing = unit(0, "lines"),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

```

### whole fig 1
```{r fig1}

layout <- "
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
#aaaaaaaaaaaaaaaaaaa
bbbbbbbbbbbbbbbbbbbb
bbbbbbbbbbbbbbbbbbbb
bbbbbbbbbbbbbbbbbbbb
"

fig1 <- article_funnel + data_collection + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 1, face = "bold"))

ggsave("figs/fig_1.wmf", 
       plot = fig1, 
       device = "wmf",
       width = fig_w, 
       height = fig_h*1.6,
       units = units)

```

## Fig 2 - Articles by country, phase, indications, route of admin

### prep

```{r fig2-prep}
# by country A
summary_country <- dat %>%
  group_by(country) %>% 
  summarise(n = length(country))

# phase B
summary_phase <- dat %>%
  group_by(phase) %>% 
  summarise(n = length(phase),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

summary_phase_condensed <- dat %>%
  group_by(phase_condensed) %>% 
  summarise(n = length(phase_condensed),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

# indication C
summary_indication <- dat %>%
  group_by(indication) %>% 
  summarise(n = length(indication),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

# route D
summary_route <- dat %>%
  group_by(route) %>% 
  summarise(n = length(route),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

```

### each panel figure 2
```{r fig2}
#by country
country_bar <- summary_country %>% 
  ggplot(aes(x = reorder(country, n), y = n)) +
  geom_bar(stat = "identity", 
           # fill = pal[4],
           fill = "black",
           size = 0, width = 0.8) +
  xlab("") +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 15, 5),
                     minor_breaks = seq(0, 15, 1),
                     limits = c(0, 16),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

# by phase
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(phase,n), y = n)) +
  geom_bar(stat = "identity", 
           # fill = pal[4],
           fill = "black",
           size = 0, 
           width = 0.8) +
  xlab("") +
 # geom_text(aes(y = n + 4, label = pc) ) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 20, 10),
                     minor_breaks = seq(0, 20, 1),
                     limits = c(0, 25),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

# by indication
indication_bar <- summary_indication %>% 
  ggplot(aes(x = reorder(indication, n), y = n)) +
  geom_bar(stat = "identity", 
           # fill = pal[4], 
           fill = "black", 
           size = 0, 
           width = 0.8) +
  xlab("") +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 12, 2),
                     minor_breaks = seq(0, 12, 2),
                     limits = c(0, 13),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

# by route
route_bar <- summary_route %>% 
  ggplot(aes(x = reorder(route, n), y = n)) +
  geom_bar(stat = "identity", 
           # fill = pal[4],
            fill = "black",
           size = 0, 
           width = 0.8) +
  xlab("") +
#  geom_text(aes(y = n + 3, label = pc) ) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 25, 5),
                     minor_breaks = seq(0, 25, 1),
                     limits = c(0, 25),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

```

### whole fig 2
```{r fig2-layout}

layout <- "
a#
a#
ad
ad
ad
ad
ad
bd
bd
bd
cd
cd
cd
cd
"

fig2 <- country_bar + phase_bar + route_bar + indication_bar +
  plot_layout(design = layout) +
  plot_annotation(caption = "Number of Articles",
                  tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.caption = element_text(size = 9, hjust = 0.5, face = "plain"))

ggsave("figs/fig_2_black.wmf", 
       plot = fig2, 
       device = "wmf",
       width = fig_w_2, 
       height = fig_h*2,
       units = units)

```

## Fig 3 - Articles by sources, allo/auto stem/stromal

### prep

```{r fig3_prep}
# source 3a
summary_source <- dat %>% 
  group_by(source) %>% 
  summarise(n = length(source),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

# allo va auto 3b
summary_allo_auto <- dat %>% 
  group_by(allo_auto) %>% 
  summarise(n = length(allo_auto),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

# stem vs stromal 3b
summary_stem_stromal <- dat %>% 
  group_by(stem_stromal) %>% 
  summarise(n = length(stem_stromal),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
```
### each panel figure 2

```{r fig_3}
# source
source_bar <- summary_source %>% 
  ggplot(aes(x = source, y = n)) +
  geom_bar(stat = "identity",
           # fill = pal[4],
           fill = "black",
           size = 0, 
           width = 0.8) +
  scale_y_continuous(name = "Number of Articles",
                    breaks = seq(0, 50, 10),
                    minor_breaks = seq(0, 50, 10),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black",
                                  angle = 285, vjust = 0.5, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())

# allo - auto
allo_auto_bar <- summary_allo_auto %>% 
  ggplot(aes(x = allo_auto, y = n)) +
  geom_bar(stat = "identity",
           # fill = pal[4], 
           fill = "black",
           size = 0, 
           width = 0.8) +
  scale_y_continuous(name = NULL,
                    breaks = seq(0, 50, 10),
                    minor_breaks = seq(0, 50, 10),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black", 
                                  angle = 285, vjust = 0.5, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())

# stem_stomal
stem_stromal_bar <- summary_stem_stromal %>%
  ggplot(aes(x = stem_stromal, y = n)) +
  geom_bar(stat = "identity",
           # fill = pal[4], 
           fill = "black",
           size = 0, 
           width = 0.8) +
  scale_y_continuous(name = NULL,
                    breaks = seq(0, 50, 10),
                    minor_breaks = seq(0, 50, 10),
                    limits = c(0, 60),
                    expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black", 
                                   angle = 285, vjust = 0.5, hjust = 0),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank())
```

### whole fig 3
```{r fig_3-layout}
layout <- "
aaabbbcccc
"

fig3 <- source_bar  + allo_auto_bar + stem_stromal_bar +
  plot_layout(design = layout) +
  plot_annotation(caption = "Source",
                  tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.caption = element_text(size = 9, hjust = 0.5, vjust = 1, face = "plain"))

ggsave("figs/fig_3_black.wmf", 
       plot = fig3, 
       device = "wmf",
       width = fig_w_1.5, 
       height = fig_h,
       units = units)
```

## Fig 4 - characterisation by article

### prep

```{r fig4_prep}
# test stringency
summary_test <- dat %>% 
  group_by(test_results) %>% 
  summarise(n = length(test_results),
            pc = paste0(round(100 * n / 84, 1 ), "%"))
# phase by stringency
summary_phase_test <- dat %>%
  group_by(phase_condensed, test_results) %>% 
  summarise(n = length(phase_condensed))

# number of markers reported in articles either reporting % by batch of average percentages (p)
summary_marker_p_count <- dat %>% 
  filter(test_results != "3. Tests done but % not reported") %>%
  filter(test_results != "4. Stated as 'standard phenotype'") %>%
  filter(test_results != "5. No characterisation discussed") %>%
  group_by(marker_p_count) %>% 
  summarise(n = length(marker_p_count))

```
### each panel figure 4

```{r fig_4}
# test stringency overall
test_bar <- summary_test %>% 
  ggplot(aes(x = test_results, y = n, fill = test_results)) +
  geom_bar(stat = "identity", size = 0) +
  xlab("") +
  geom_text(aes(y = n + 6, label = pc), size = 3 ) +
  scale_fill_manual(values = rev(pal[1:5]),
                    name = "Test results") +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 50, 5),
                     minor_breaks = seq(0, 50, 1),
                     limits = c(0, 53),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), ) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 30),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none") +
  coord_flip()

# tests stringency by phase
phase_test <- summary_phase_test %>% 
  ggplot(aes(x = phase_condensed, y = n, fill = test_results)) +
  geom_bar(stat = "identity", size = 0, width = 0.8) +
  scale_fill_manual(values = pal,
                    name = NULL) +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 55, 10),
                     minor_breaks = seq(0, 55, 5),
                     limits = c(0, 55),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "Phase") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 0, l = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none")

# 
markerp_bar <- summary_marker_p_count %>% 
  ggplot(aes(x = marker_p_count, y = n)) +
  geom_bar(stat = "identity", fill = "black", size = 0, width = 0.8) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 12, 2),
                     minor_breaks = seq(0, 12, 2),
                     limits = c(0, 12),
                     expand = c(0, 0)) +
  scale_x_continuous(name = "Number of markers",
                     breaks = seq(0, 11, 1),
                    minor_breaks = seq(0, 12, 2),
                    limits = c(0.5, 12),
                     expand = c(0, 0)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.margin = margin(t = 0, r = 0, l = 0))

```

### whole fig 4
```{r fig_4-layout}

layout <- "
aaaaaabbbcccc
"

fig4 <- test_bar + phase_test + markerp_bar +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))

ggsave("figs/fig_4.wmf", 
       plot = fig4, 
       device = "wmf",
       width = fig_w_2, 
       height = 2.2,
       units = units)
```

## Fig 5 - Characterisation detail

### prep

```{r}

# select the test columns and count the number of articles for each marker and category
test_cols <- names(select(dat, ends_with("_test")))

test_by_attribute <- dat %>% 
  select(test_cols, ref_number) %>% 
  pivot_longer(cols = -ref_number, 
               names_to = "marker", 
               values_to = "status")

# process the names to remove _test

test_by_attribute <- test_by_attribute %>% 
  mutate(marker2 = str_replace(marker, "_test", ""))

# change levels of status and markers.
# markers
# process test_col names
test_cols <- str_replace(test_cols, "_test", "")
# set the levels in that order
test_by_attribute$marker2 <- factor(test_by_attribute$marker2, 
                                    levels = test_cols)
test_by_attribute$ref_number <- factor(test_by_attribute$ref_number)
# status
# recode status
test_by_attribute <- test_by_attribute %>%
  mutate(test_status = recode_factor(status,
                                     p = "Performed, value reported (by batch or average)",
                                     m = "Performed, no value reported",
                                     n = "Not performed"))

test_by_attribute$test_status <- fct_relevel(test_by_attribute$test_status,
                                             "Not performed",
                                             "Performed, no value reported",
                                             "Performed, value reported (by batch or average)")
test_by_attribute <- test_by_attribute %>% 
  select(ref_number, marker2, test_status)

# test_by_attribute$n <- 1

# create the summary row & column
# test_by_attribute_aug <- bind_rows(
#   test_by_attribute,
#   test_by_attribute %>% 
#     filter(test_status == "Performed, value reported (by batch or average)") %>% 
#     group_by(ref_number, .drop = FALSE) %>% 
#     summarize(
#       n = sum(n),
#       marker2 = "Total P",
#       test_status = "Performed, value reported (by batch or average)"
  #   ),
  # test_by_attribute %>% 
  #   filter(test_status == "Performed, value reported (by batch or average)") %>% 
  #   group_by(marker2, .drop = FALSE) %>% 
  #   summarize(
  #     n = sum(n),
  #     ref_number = "Total P",
  #     test_status = "Performed, value reported (by batch or average)"
#     )
# )
# 
# test_by_attribute_aug <- bind_rows(
#   test_by_attribute_aug,
#   test_by_attribute %>% 
#     filter(test_status == "Performed, no value reported") %>% 
#     group_by(ref_number, .drop = FALSE) %>% 
#     summarize(
#       n = sum(n),
#       marker2 = "Total M",
#       test_status = "Performed, no value reported"
  #   ),
  # test_by_attribute %>% 
  #   filter(test_status == "Performed, no value reported") %>% 
  #   group_by(marker2, .drop = FALSE) %>% 
  #   summarize(
  #     n = sum(n),
  #     ref_number = "Total M",
  #     test_status = "Performed, no value reported"
#     )
# )

# test_by_attribute_aug$marker2 <- factor(test_by_attribute_aug$marker2, 
#                                     levels = c(test_cols, "Total P", "Total M"))
# test_by_attribute_aug$ref_number <- factor(test_by_attribute_aug$ref_number, 
#                                     levels = levels(test_by_attribute$ref_number))

test_by_attribute_summary <- test_by_attribute %>% 
  group_by(marker2, test_status) %>% 
  summarise(n = length(marker2))

isct_pos <- c("CD73", "CD90", "CD105")
isct_neg <- c("CD34", "CD45", "CD11b", "CD14", "CD79a", "CD19", "HLA-DR")
isct_not <- c("viability",  "CD3" ,"CD13", "CD31", "CD133", "CD166", "CD29", "CD44", "CD146", "CD271", "STRO-1", "MSCA-1", "SSEA-4")

test_by_attribute <- test_by_attribute %>%
    mutate(ISCT = case_when(marker2 %in% isct_pos ~ "+",
                            marker2 %in% isct_neg ~ "–",
                            marker2 %in% isct_not ~ ""))
test_by_attribute_summary <- test_by_attribute_summary %>%
    mutate(ISCT = case_when(marker2 %in% isct_pos ~ "+",
                            marker2 %in% isct_neg ~ "–",
                            marker2 %in% isct_not ~ ""))
codes <- data.frame(marker2 = test_cols)
codes <- codes %>%
  mutate(ISCT = case_when(marker2 %in% isct_pos ~ "+",
                          marker2 %in% isct_neg ~ "–",
                          marker2 %in% isct_not ~ ""))
test_by_attribute$ref_number <- as.numeric(as.character(test_by_attribute$ref_number))
```

### each panel in figure 5

```{r}

# create the plot
# fig5 <- test_by_attribute_aug %>% 
#   ggplot() + 
#   geom_tile(aes(x = marker2, y = ref_number, fill = test_status), colour = "white") + 
#   scale_fill_manual(values =  c("#BABABA", pal[c(3, 5)]),
#                     name = "Test results",
#                     guide = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5 )) +
#   scale_x_discrete(name = "Characterisation attribute",
#                      expand = c(0.02, 0)) +
#   scale_y_discrete(expand = c(0, 0), name = "Reference") +
#   geom_text(data = filter(test_by_attribute_aug, marker2 == 'Total P' | ref_number == 'Total P'| marker2 == 'Total M' | ref_number == 'Total M'), 
#              aes(x = marker2, y = ref_number, label = n), size = 3, face = "bold") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(size = 8, colour = "black",
#                                    angle = 300, hjust = 0, vjust = 0.3),
#         axis.text.y  = element_text(size = 8, colour = "black"),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         legend.position = "top",
#         legend.text = element_text(size = 8, colour = "black"),
#         legend.key.size = unit(0.2, "in"))

# fig5 <- test_by_attribute_aug %>% 
#   ggplot() + 
#   geom_tile(aes(x = ref_number, y = marker2, fill = test_status), colour = "white") + 
#   scale_fill_manual(values =  c("#BABABA", pal[c(3, 5)]),
#                     name = "Test results",
                    # guide = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5 )) +
#   scale_x_discrete(name = "Reference",
#                      expand = c(0.02, 0)) +
#   scale_y_discrete(expand = c(0, 0), name = "Characterisation attribute" ) +
#   geom_text(data = filter(test_by_attribute_aug, marker2 == 'Total P' | ref_number == 'Total P'| marker2 == 'Total M' | ref_number == 'Total M'), 
#              aes(x = ref_number, y = marker2,  label = n), size = 3, face = "bold") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(size = 8, colour = "black",
#                                    angle = 300, hjust = 0, vjust = 0.3),
#         axis.text.y  = element_text(size = 8, colour = "black"),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         legend.position = "top",
#         legend.text = element_text(size = 8, colour = "black"),
#         legend.key.size = unit(0.2, "in"))

fig5a <- test_by_attribute %>% 
  ggplot() + 
  geom_tile(aes(x = ref_number, y = marker2, fill = test_status), colour = "white") + 
  scale_fill_manual(values =  c("#BABABA", pal[c(3, 5)]),
                    name = "Test results",
                    guide = guide_legend(nrow = 3)) +
  scale_x_continuous(name = "Reference",
                     breaks = seq(0, 84, 5),
                     minor_breaks = seq(0, 84, 5),
                     limits = c(1, 87),
                     expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0), name = "Characterisation attribute") +
  # geom_text(aes(x = 85, y = marker2 , label = ISCT),
  #           size = 2,
  #           face = "bold") +
  # coord_cartesian(xlim = c(0, 84), # This focuses the x-axis on the range of interest
  #                 clip = 'off') +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8, colour = "black", hjust = 0.5,),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
         legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in"))

fig5b <- test_by_attribute_summary %>% 
  ggplot() +
  geom_bar(aes(x = n, y = marker2, fill = test_status),
           stat = "identity", width = 0.8, size = 0) +
  scale_fill_manual(values =  c("#BABABA", pal[c(3, 5)])) +
  scale_x_continuous(name = "Number of Articles",
                     breaks = seq(0, 90, 10),
                     minor_breaks = seq(0, 90, 10),
                     limits = c(0, 85),
                     expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0.5), name = NULL, labels = codes$ISCT) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 10, colour = "black",
                                    face = "bold", vjust = 0.3),
        axis.ticks.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none")

```

### the whole figure 5

```{r}

layout <- "
aaab
"

fig5 <- fig5a + fig5b &
  plot_layout(design = layout) &
  plot_annotation(tag_levels = "A") &
  geom_hline(yintercept = 1.5) &
   geom_hline(yintercept = 11.5) &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.margin = margin(0,0,0,0))

ggsave("figs/fig_5.wmf", 
       plot = fig5, 
       device = "wmf",
       width = fig_w_2, 
       height = fig_h * 1.2,
       units = units)
```

## Fig 6 - functionality assessments

### prep

```{r}
# panel a
functionality <- dat %>% 
  select(O, A, C, other) %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "status") %>% 
  group_by(type, status) %>% 
  summarise(n = length(status),
            pc = paste0(round(100 * n / 84, 1 ), "%")) %>% 
  ungroup(type, status)

# status
# relevel status
functionality$status <- fct_relevel(functionality$status,
                                    "Not mentioned",
                                    "Performed, no value reported",
                                    "Performed, value reported")

# recode functionality
functionality <- functionality %>%
  mutate(type = recode_factor(type,
                              other = "Other functionality",
                              C = "Chondrogenesis",
                              A = "Adipogenesis",
                              O = "Osteogenesis"))

functionality$type <- fct_relevel(functionality$type,
                                         "Other functionality",
                                         "Chondrogenesis",
                                         "Adipogenesis",
                                         "Osteogenesis")

# panel b
moa_isct_summary <- dat %>% 
  group_by(MOA, stem_stromal) %>% 
  summarise(n = length(MOA),
            pc = paste0(round(100 * n / 84, 1 ), "%"))

moa_isct_summary$stem_stromal <- fct_relevel(moa_isct_summary$stem_stromal,
                                         "Stem",
                                         "Stromal",
                                         "Multipotent Stromal",
                                         "Regenerative")

moa_isct_summary$MOA <- fct_relevel(moa_isct_summary$MOA,
                                    "Not stated",
                                    "Multiple",
                                    "Differentiation",
                                    "Immune",
                                    "Paracrine")

# panel c
diffmoasum <- dat %>% group_by(MOA, differentiation) %>% 
  summarise(n = length(differentiation))

diffmoasum$MOA <- fct_relevel(diffmoasum$MOA,
                                     "Not stated",
                                    "Multiple",
                                    "Differentiation",
                                    "Immune",
                                    "Paracrine")

diffmoasum$differentiation <- fct_relevel(diffmoasum$differentiation,
                                          "No",
                                          "Yes")

```

### each panel figure 6

```{r}
# panel a
fig6a <- functionality %>% 
  ggplot() +
  geom_bar(aes(x = type, y = n, fill = status, label = pc),
           stat = "identity", size = 0) +
  scale_fill_manual(values =  c("#BABABA", pal[c(3, 5)]),
                    name = " Functionality",
                    guide = guide_legend(nrow = 4,
                                         title.position = "top",
                                         title.hjust = 0 )) +
  scale_y_continuous(name = "Number of Articles",
                     breaks = seq(0, 80, 10),
                     minor_breaks = seq(0, 80, 10),
                     limits = c(0, 85),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0.5), name = NULL) +
  geom_text(data = filter(functionality, 
                          status == "Not mentioned"),
            aes(x = type, 50, label = pc) ) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 30),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in")) +
  coord_flip()

# panel b
fig6b <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal, 
                    name = "Nomenclature",
                    guide = guide_legend(nrow = 4,
                                         title.position = "top",
                                         title.hjust = 0 )) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 30, 5),
                     minor_breaks = seq(0, 30, 5),
                     limits = c(0, 35),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "MOA") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 0, l = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in")) +
  coord_flip()

# panel c
fig6c <- diffmoasum %>%
  ggplot(aes(x = MOA, y = n, fill = differentiation)) +
  geom_bar(stat = "identity", size = 0) +
   scale_fill_manual(values = c("#BABABA", pal[5]), 
                    name = "Differentiation capacity",
                    labels = c("Not demonstrated", "Demonstrated"),
                    guide = guide_legend(nrow = 4, 
                                         title.position = "top",
                                         title.hjust = 0)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 30, 5),
                     minor_breaks = seq(0, 30, 5),
                     limits = c(0, 35),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL, labels = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 0, l = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in")) +
  coord_flip()

```

### the whole figure 6

```{r}

layout <- "
abc
"

fig6 <- fig6a + fig6b + fig6c +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))
ggsave("figs/fig_6.wmf", 
       plot = fig6, 
       device = "wmf",
       width = fig_w_2, 
       height = 3,
       units = units)

```

## relationship between phase_condensed and test_results

```{r}
res <- table(dat$phase_condensed, dat$test_results) %>%
  fisher.test(simulate.p.value = TRUE)
test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))

```

## Relationship between MOA and stem/stromal

```{r}
res <- table(dat$stem_stromal, dat$MOA) %>%
  fisher.test(simulate.p.value = TRUE)
test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))

# without the regen and multi cats
dat_sub <- dat %>%
  filter(stem_stromal != "Regenerative") %>% 
  filter(stem_stromal != "Multipotent Stromal") %>% droplevels()
res <- table(dat_sub$stem_stromal, dat_sub$MOA) %>% 
  fisher.test(simulate.p.value = TRUE)

```

## Figure 10 assocation between differentiation and MOA

```{r}
res <- table(dat$differentiation, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)

```
# CD values figures

```{r}
cd_dat_summary <- cd_dat %>% 
  group_by(Ref, marker) %>% 
  summarise(mean_value = mean(value),
            n = length(value),
            se_value = sd(value)/sqrt(n))
str(cd_dat_summary)
# length(unique(cd_dat$marker)) *length(unique(cd_dat$Ref))

# bind the range and the sinlge value dataframes
cd_dat_summary <- rbind(data.frame(cd_dat_summary), cd_dat2)

```

```{r}
# count the number of refs that have values
cd_refs_by_marker <- cd_dat_summary %>% 
  group_by(marker) %>%
  summarise(n_refs = length(mean_value[!is.na(mean_value)]))

cd_refs_by_marker$marker
isct <- c("Viability","CD73", "CD90", "CD105", "CD34", "CD45" , "CD11b", "CD14" , "CD79α", "CD19", "HLA-DR")
cd_refs_by_marker$ISCT <- "Other marker"
cd_refs_by_marker$ISCT[cd_refs_by_marker$marker %in% isct] <- "ISCT phenotype marker"

cd_refs_by_marker$ISCT <- fct_relevel(cd_refs_by_marker$ISCT,
                              "ISCT phenotype marker",
                              "Other marker")

```

```{r cd-refs-bar}
cd_refs_bar <- cd_refs_by_marker %>% 
  arrange(ISCT, n_refs) %>%
         mutate(marker = factor(marker, levels = marker) ) %>% 
  ggplot(aes(x = reorder(marker, ISCT), y = n_refs, fill = ISCT )) +
  geom_bar(stat = "identity",  size = 0) +
  xlab("") +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 15, 1),
                     minor_breaks = seq(0, 15, 1),
                     limits = c(0, 15),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_manual(values = pal[c(5, 2)], 
                    name = NULL) +
  labs(title = "Wilson, Alison, Figure X3. TOP" ) +
  annotate("segment", x = 0.5, xend = 10.5, 
           y = 14, yend = 14,
           colour = "black", size = 1) +
  annotate("segment", x = 0.58, xend = 0.58, 
           y = 14, yend = 13.5,
           colour = "black", size = 1) +
  annotate("segment", x = 10.38, xend = 10.38, 
           y = 14, yend = 13.5,
           colour = "black", size = 1) +
  annotate("text", x = 3,  y = 10, 
           label = "ISCT phenotype\nmarker", size = 4) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

ggsave("figs/fig_x3_cd_refs_bar.wmf", 
       plot = cd_refs_bar, 
       device = "wmf",
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}

cd_dat_summary_values <- cd_dat_summary %>% 
  filter(!is.na(mean_value))

cd_dat_summary_values$ISCT <- "Other marker"
cd_dat_summary_values$ISCT[cd_dat_summary_values$marker %in% isct] <- "ISCT phenotype marker"

cd_dat_summary_values$ISCT <- fct_relevel(cd_dat_summary_values$ISCT,
                              "ISCT phenotype marker",
                              "Other marker")
```

```{r}
cd_values_point_isct <- cd_dat_summary_values %>% 
  filter(ISCT == "ISCT phenotype marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x4_cd_values_isct.wmf", 
       plot = cd_values_point_isct, 
       device = "wmf",
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}
cd_values_point_other <- cd_dat_summary_values %>% 
  filter(ISCT == "Other marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x5_cd_values_isct.wmf", 
       plot = cd_values_point_other, 
       device = "wmf",
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

```

```{r}

# disease targets
dat %>% group_by(Indication) %>% 
  summarise(n = length(Indication)) %>% 
  ggplot(aes(x = Indication, y = n)) +
  geom_bar(stat="identity")

```